---
title: "Exploring Ami Bennett’s browsing data"
author: "David Duncan"
date: "19/09/2017"
output: html_document
---

# Setup the `R` session
Load packages and set document defaults
```{r setup, include=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(readxl, tidyverse, ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

# Prepare the data

Considerable pre-prep was done Excel-side by Ami, and myself, including coding up the status info with respect to browsing evidence.

Data prep now occurs in `prepareData.R`, so we'll just read them in here with the following.
```{r}
load("browseDFs.RData")
```

## Seedling states and fates
This factor required a bit more work to make sure it graphs in the right order and with labels that are easy to interpret.
```{r ordinal factors}
saplings$state <- factor(saplings$state, levels = c("A",  "A_wk", "A_b", "A_wk_b", "A_B", "D", "D_b", "D_B", "D_nt"), ordered = TRUE)

```

## Group and summarise observations for different levels

### Summary each sapling of final state

### Tracking through experiment at site level

```{r survival summary per site}
survival_site <- saplings %>% 
  group_by(site, context, Treat, Time) %>% 
  summarise(total = n(),
            nDead = sum(Dead),
            propAlive = 1- (nDead/total),
            days = mean(elapsedDays)) %>%
  mutate(link = paste0(site, ".", Time))
```


### Version at level of site context

```{r for contexts}
survival_context <- saplings %>% 
  group_by(context, Treat, Time) %>% 
  summarise(total = n(),
    nDead = sum(Dead),
    propAlive = 1- (nDead/total),
    days = mean(elapsedDays)
  )
```


```{r make plot label and colour vectors, echo=FALSE}
contextNames <- c(
  'Buloke woodland' = "Buloke woodland",
  'Mallee' = "Adjacent mallee woodland",
  'Open grassland' = "Open grassland",
  'Wattle dune' = "Adjacent wattle dune"
)

contextNameNarrow <- c(
  'Buloke woodland' = "Within buloke\nstand",
  'Mallee' = "Adj. mallee\nwoodland",
  'Open grassland' = "Open\ngrassland",
  'Wattle dune' = "Adj. wattle\ndune"
)

treatName <- c(
  'Open' = "Open to\nherbivores",
  'Partial' = "Partial\nexclosure",
  'Total' = "Full\nexclosure"
  )

treatNameFactSheet <- c(
  'Open' = "Unguarded",
  'Partial' = "Partially\nguarded",
  'Total' = "Full guards"
  )

```

# Seedling survival

## plot survival over the course of the experiment so far
Overall result, proportion of seedlings surviving as a function of the location context.
```{r survival by treatment with days accumulated, echo=FALSE}
ggplot(data = survival_context, mapping = aes(x = days, y = propAlive, color = Treat)) + 
  geom_point(size=2) + geom_line(lwd=1) +
  facet_grid(~context, labeller = as_labeller(contextNameNarrow)) +
  theme(strip.text = element_text(size = 10)) +
  labs(y = "Proportion of seedlings still alive", x = "Days since planting (Spring '16)", colour = "Treament")
```

## Break down to site level to look at consistency

```{r survival by site with days accumulated, echo=FALSE}
ggplot(data = survival_site, mapping = aes(x = days, y = propAlive, color = Treat)) + 
  geom_point(size=2) + geom_line(aes(group=site)) +
  facet_grid(Treat~context) + 
  labs(y = "Proportion of seedlings alive", x = "Days since planting (Spring '16)")
```

Another helpful view of that might be to contrast treatments at each site as small multiples.
```{r}
ggplot(survival_site, aes(x = days, y = propAlive, color = Treat)) +
  geom_point(size=2) + geom_line(aes(group=Treat)) +
  #geom_linerange(aes(ymin = lwr_prBrowsed, ymax = upp_prBrowsed)) +
  facet_wrap(context~site) + 
  labs(y = "Proportion of seedlings browsed", x = "Days since planting (Spring '16)")
```

## plot status summaries over the course of the experiment so far

### Summary of state

```{r plot sapling state per visit}
stateCols <- c("A" = "chartreuse4", "A_wk" = "yellowgreen", "A_b" = "yellow3", "A_wk_b" = "cornsilk2", "A_B" = "pink", "D" = "black", "D_b" = "red1", "D_B" = "red3", "D_nt" = "grey")

saplings %>% 
  count(context, Treat, Time, state) %>% 
ggplot(mapping = aes(x = Time, y = n)) + 
  geom_bar(aes(fill = state), stat = "identity", position = position_fill(reverse = TRUE) ) + 
  scale_fill_manual(values = stateCols) +
  facet_grid(context~Treat, labeller = labeller(context = as_labeller(contextNameNarrow), Treat = as_labeller(treatName))) + 
  labs(y = "Proportion of sample", x = "Visit number") 
```

```{r plot sapling state at final visit}
stateCols2 <- c("A" = "chartreuse4", "A_d" = "yellow3", "D" = "black", "D_d" = "red3", "D_B" = "red1", "D_nt" = "grey")

stateCols_bw <- c("Alive no damage" = "black", "Alive + damaged" = "grey25",
                  "Alive + extreme damage" = "pink", 
                  "Dead no damage" = "grey85", 
                  "Dead + damaged" = "grey50", 
                  "Dead + extreme damage" = "red1", 
                  "Dead (no trace)" = "white")

library(dichromat)
scheme <- colorschemes$SteppedSequential.5
stateCols_dcmt <- c("Alive\nno damage" = scheme[15], "Alive\n+ damaged" = scheme[13],
                  "Alive\n+ extreme damage" = scheme[11], 
                  "Dead\nno damage" = scheme[10], 
                  "Dead\n+ damaged" = scheme[8], 
                  "Dead\n+ extreme damage" = scheme[6], 
                  "Dead\n(no trace)" = scheme[6])

border_col <- c("Alive\nno damage" = element_blank(), "Alive\n+ damaged" = element_blank(),
                  "Alive\n+ extreme damage" = element_blank(), 
                  "Dead\nno damage" = element_blank(), 
                  "Dead\n+ damaged" = element_blank(), 
                  "Dead\n+ extreme damage" = element_blank(), 
                  "Dead\n(no trace)" = "grey50")



a <- saplings %>% 
  filter(Time == 6) %>% 
  mutate(context = factor(context, 
                          levels = c("Buloke woodland", "Wattle dune",
                                     "Mallee",
                                     "Open grassland"),
                          labels = c("BW", "WD", "MW", "OG"),
                          ordered = TRUE),
         state = factor(state, 
                        levels = c("A",  "A_wk", "A_b", "A_wk_b", "A_B", "D_B","D_nt", "D_b",  "D"),
                        labels = c("Alive\nno damage",  "Alive\nno damage", "Alive\n+ damaged", "Alive\n+ damaged", "Alive\n+ extreme damage", "Dead\n+ extreme damage", "Dead\n(no trace)", "Dead\n+ damaged",   "Dead\nno damage"), ordered = TRUE)) %>% 
  count(context, Treat, Time, state) %>% 
ggplot(mapping = aes(x = context, y = n)) + 
  geom_bar(aes(fill = state), color = "black", stat = "identity", position = position_fill(reverse = TRUE) ) + 
  scale_fill_manual( "Status", values = stateCols_dcmt, guide = guide_legend(reverse=TRUE)) +
  #scale_color_manual(values = border_col) +
  facet_grid(.~Treat, labeller = as_labeller(treatName)) + 
  labs(y = "Proportion of sample", x = "Planting context")

pdf("gfx/ms_fig2.pdf", width = 6.14, height = 5.87)
print(
  a + theme(
  legend.spacing.y = unit(15, units = "pt"),
  #plot.title = element_text(face = "bold", size = rel(1.2),
  #                          hjust = 0.5),
  #       text = element_text(),
  panel.background = element_rect(colour = NA),
  plot.background = element_rect(colour = NA),
  panel.border = element_rect(colour = NA, fill = NA),
  axis.title = element_text(size = rel(1)),
  axis.title.y = element_text(vjust =2),
  #        axis.title.x = element_text(vjust = -0.2),
  axis.text.x = element_text(angle = 45, hjust = .9),
  axis.line = element_line(colour="black", size = 0.5),
  axis.ticks = element_line(),
  panel.grid.major = element_line(colour="#f0f0f0"),
  panel.grid.minor = element_blank(),
  legend.key = element_rect(colour = NA),
  #legend.spacing.y = unit(15, units = "pt"),
  legend.key.size= unit(0.2, "cm"),
  legend.key.width = unit( 0.4, "cm"),
  legend.margin = margin(0,1,0,0),
  legend.title = element_text(face="italic"),
  plot.margin=unit(c(.2,.5,1.5,1.5),"mm"),
  strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
  strip.text = element_text(face="bold"))
)
dev.off()
```





```{r}
saplings %>% 
  filter(Time == 6) %>% 
  mutate(context = factor(context, 
                          levels = c("Buloke woodland", "Wattle dune",
                                     "Mallee",
                                     "Open grassland"),
                          labels = c("Bw\nn=175", "Wd\nn=75", "Mw\nn=75", "OG\nn=150"),
                          ordered = TRUE),
         state = factor(state, 
                        levels = c("A",  "A_wk", "A_b", "A_wk_b", "D_b", "D_B", "D", "D_nt"),
                        labels = c("Alive no damage",  "Alive no damage", "Alive + damaged", "Alive + damaged","Dead + damaged", "Dead + extreme damage",  "Dead no damage", "Dead (no trace)"), ordered = TRUE)) %>% 
  count(context, Treat, Time, state) %>% 
  ggplot(mapping = aes(x = context, y = n)) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = rel(1.2), hjust = 0.5),
    text = element_text(),
    panel.background = element_rect(colour = NA),
    plot.background = element_rect(colour = NA),
    panel.border = element_rect(colour = NA, fill = NA),
    axis.title = element_text(face = "bold",size = rel(1)),
    axis.title.y = element_text(angle=90,vjust =2),
    axis.title.x = element_text(vjust = -0.2),
    axis.text = element_text(), 
    axis.line = element_line(colour="black"),
    axis.ticks = element_line(),
    panel.grid.major = element_line(colour="#f0f0f0"),
    panel.grid.minor = element_blank(),
    legend.key = element_rect(colour = NA),
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.size= unit(0.2, "cm"),
    legend.key.width = unit( 0.4, "cm"),
    legend.margin = margin(0,0,0,0),
    legend.title = element_text(face="italic"),
    plot.margin=unit(c(10,5,5,5),"mm"),
    strip.background=element_rect(colour="#f0f0f0",
                                  fill="#f0f0f0"),
    strip.text = element_text(face="bold")
    ) +
  geom_bar(aes(fill = state), stat = "identity", position = position_fill(reverse = TRUE) ) + 
  scale_fill_manual( "Status", values = stateCols_bw) +
  facet_grid(.~Treat, labeller = as_labeller(treatName)) + 
  labs(y = "Proportion of sample", x = "Planting context") 
```



# Seedling by site history check up
These graphs can be found in the "seedlingHistories" subfolder. Their primary purpose was just to check that thenested IF THEN statements executed in MS Excel hadn't produced perverse outcomes; seedlings dying and coming back to life, being fine then severely browsed then fine again, etc.

```{r Seedling lifeline plots per site, echo=FALSE, eval=FALSE}
labeller <- levels(saplings$state)

for(s in unique(saplings$site)) {
  
  checkSite <- saplings[saplings$site == s, ]
  
  png(paste("seedlingHistories/T6_site_",s,".png",sep = ""), width = 1100, height = 900)
  print(ggplot(checkSite,aes(x=Time, y = as.numeric(state))) + 
          geom_line(color="grey") + 
          geom_point(aes(colour = state, shape = factor(Dead)), size=2) +
          scale_shape_manual(values = c(19, 8)) +
          scale_color_manual(values = stateCols) +
          labs(title = paste("Site: ",s), y = "Seedling status", x = "Visit #") +
          scale_y_continuous(breaks = 1:9, labels = labeller) +
          facet_wrap(~uniqID)
  )
  
  dev.off()
  
}
```

I have also done the same for the growth data over the course of the experiment.
```{r make sapling growth summaries for each site and save for ron, eval=FALSE}
for(s in unique(saplings$site)) {
  
  checkSite <- saplings[saplings$site == s, ]
  
  png(paste("seedlingHistories/growth_T6_Site_",s,".png",sep = ""), width = 1100, height = 900)
  print(ggplot(checkSite[!is.na(checkSite$HtCM),],aes(x=Time, y = HtCM)) +  # note the call to !is.na() so that lines join across NA gaps.
          geom_line(color="grey") + 
          geom_point(aes(colour = state, shape = factor(Dead)), size=2) +
          scale_shape_manual(values = c(19, 8)) +
          scale_color_manual(values = stateCols) +
          labs(title = paste("Site: ",s), y = "Seedling Height (CM)", x = "Visit #") +
          facet_wrap(~uniqID, scales = "free_y")
  )
  
  dev.off()
  
}
```


```{r look at growth summaries}
ggplot(saplings_summary, mapping = aes(x = ht_change, fill = Treat)) + geom_histogram() + facet_grid(Treat~context)
```


```{r}
ggplot(saplings_summary, aes(x=start_ht, y= end_ht, color=Treat)) + geom_point() + facet_wrap(~site)
```


# Browser pellet data

```{r pellet summaries, echo=FALSE}

load("pellets.Rdata")

ggplot(data = plt_plot, 
       mapping = aes(x = factor(Time), y = nPellets+1, color =browser)) +
  geom_boxplot() + 
  #geom_jitter() +
  #geom_line() + 
  facet_grid(context~browser) + scale_y_log10()

ggplot(data = plt_plot, 
       mapping = aes(x = factor(Time), y = rPellets+1, color = browser)) +
  geom_boxplot() + 
  #geom_jitter() +
  #geom_line() + 
  facet_grid(context~browser) + scale_y_log10()

```

# Accumulation rate view ----
```{r Accumulation rate view, eval = FALSE, echo=FALSE}
ggplot(data = plts, mapping = aes(x = dysAcc, y = rPellets+0.00005)) + 
  geom_jitter(aes(size = nPellets, color = browser), width = 1.5) + 
  labs(x = "Days accumulated before count", y = "Calculated rate of pellets / day (log scale)") +
  scale_y_log10() + facet_wrap(context~site)
```


```{r Pellet rate by visit, eval = FALSE, echo=FALSE}
ggplot(data = plts, 
       mapping = aes(x = factor(Time), y = rPellets+0.0005, color = browser)) + 
  geom_point(position = "jitter") +
  facet_wrap(context~site) + scale_y_log10()
```

# Influence of observer by site by time? ----
In the field we worried about the possibility of observer differences in the number of pellets observed. Here's what the data looks like.
```{r Observer by site by time, eval=FALSE}
ggplot(data = plts, mapping = aes(x = factor(Time), y = nPellets+0.5, 
                                  fill = obs)) + 
  geom_boxplot() + 
  facet_wrap(context~site) + 
  scale_y_log10()
```
**I turned this one off (eval = FALSE)**

# Browser activity versus browsing and mortality 
Is activity as represented by dung correlated with browsing and death due to browsing?

```{r, echo=FALSE}
survival_site <- left_join(survival_site, 
                           select(ungroup(plt_plot), 
                                  c(link, browser, nPellets)), 
                           by = "link")

survival_site %>% filter(browser == 'Lagomrph') %>% 
ggplot(aes(x = nPellets+0.1, y = propAlive, color = Treat)) + 
  geom_point() + #geom_smooth(method = "gam") + 
  facet_wrap(~Time) + labs(x = 'number of lagomorph pellets') +
  scale_x_log10()
```


