---
title: "RPP 1819 P12: Breaking Buloke Woodlands out of the Browse trap"
subtitle: "Report to Parks Victoria"
author: "David Duncan and Ami Bennett"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    prettydoc::html_pretty:
        theme: cayman
        highlight: github
bibliography: refs.bib  
csl: AJBish.csl 
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, results='hide', comment=NA, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# packages ----
library(dplyr)
library(openxlsx)
library(tidyr)
library(ggplot2)

path_to_data <- "data/NESP1_2_2_buloke_survival_23_dec2020.xlsx"

```

# Executive Summary

This project is investigating the potential to continue scaled-down
monitoring of the growth and survival of hand-planted buloke
(*Allocasuarina luehmannii*) seedlings at Wyperfeld National Park.

The proposal addresses key questions (numbered 3 and 4) identified by Hauser and Lahoz-Monfort [-@hauser2014] in their analysis of Adaptive Management of Grazing Pressure on Pine-Buloke Vegetation at Wyperfeld National Park. Specifically:

3.	Under what conditions (fire, rainfall, grazing) can Pine and Buloke seedlings escape grazing pressure?
4.	Under what conditions (fire, rainfall, grazing, age and size) can Pine and Buloke individuals reach reproductive maturity?

We will generate invaluable data on seedling survival and growth required to validate a survival model built on data from Bennett et al, but also to develop estimates of time required for seedlings to escape from vulnerability to browsing damage, sometimes called ‘escape height’.

This project builds on the successful large scale buloke seedling survival experiment at Wyperfeld National Park [@bennett2020]. The objectives of this phase of the project are to:

1.	Identify patterns of established seedling survival, browsing damage, and growth rates as a function of landscape position and herbivore access.
2.	Determine if a cheaper measure of herbivore activity can replace the measure used in the first phase of the survival experiment . 
3.	Report on the suitability of the various methods for future field experiments or management applications where indices of herbivore activity are required.

The most recently funded phase of the project ran to June 2020, however
due to the chaotic situation of the Covid19 pandemic, no final report
was delivered.



## Outputs to date

@bennett2020

An accessible guide to restoration of Buloke Woodlands in the semi-arid part of their range [@duncan2021] (defined that way given that Macaulay's [-@macaulay2003] guide remains an excellent reference for the more southern part of the distribution.

## Evolution of field experiment

We made a number of important changes to the field experiment.

2019 -

2020 -

# Method

```{r}
load("data/browseDFs.Rdata")
load("data/herb_activity.Rdata")


```


```{r extract calibration subsets}
accumm_calib <- saplings_surv %>%
  group_by(site, Time, Treat, context, link) %>% 
  filter(Time == 6) %>% 
  summarise(
    accumm_days = first(days),
    lgm_15.75m2 = first(Lagomrph_site_mean),
    roo_15.75m2 = first(Roo_site_mean),
    lgm_0.1m2_yr = first(m_nP_0.1m2_yr),
    lgm_0.1m2_yr_2 = ((lgm_15.75m2/157.5)/accumm_days)*365,
    roo_0.1m2_yr = ((roo_15.75m2/157.5)/accumm_days)*365
  )
```



# Results


```{r}
plt_compare <- ggplot(link_plt, aes(Time, d_rab_mtz)) + 
  geom_point() + 
  facet_wrap(~site)
```



# Discussion


# Outlook

Matt Whitney MSc (supported by PV, Bill Borthwick, Botany Foundation and )

Review paper

Rabbit model

GA discussion, referred by Brendan Rodgers

# References
